CO 789, Homework 1

### 2024-01-30 co789/homework-01/q03/main.tex
Office hour notes

\subsection*{(2)}
I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in. Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$. Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
https://www.overleaf.com/project/65a54df4dabb2a4ef8fbbc0c
The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:

@@

\subsection*{(4)}
\subsubsection*{(a)}
A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.

>>>>

\subsection*{(2)}
I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in. Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$. Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.

The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:

@@

\subsection*{(4)}
\subsubsection*{(a)}
A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow (\mathbf{b} - A\hat{\mathbf{s}}) \mod q$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.

### 2024-01-30 co789/homework-01/q05/main.tex
Office hour notes

\subsection*{(2)}
For the remainder of this problem, we assume that full-rank LWE with parameters $(m, n, q, U_s, \chi_e)$ exist, which means that $n \leq m$.

@@

we can discard the linearly dependent columns of $A$ and feed the truncated sample $(A_1, \mathbf{b})$ into the full-rank LWE oracle. If the corresponding full-rank Search-LWE has unique solution denoted by $\mathbf{s}^\prime$, then it must be that $\mathbf{b} - A_1\mathbf{s}^\prime = \mathbf{e}$, where $\mathbf{e}$ is exactly the error term from the original Search-LWE instance $(A, \mathbf{b})$.

With the error term recovered, the Search-LWE instance becomes solving noiseless linear equations. Because $A$ is not necessarily full-rank, there may be more than one solutions, but since the secret is uniformly randomly sampled, each solution is equally likely the true secret. I claim that this is the best we can do in terms of solving Search-LWE for non-full-rank $A$.

>>>>

\subsection*{(2)}
For the remainder of this problem, we assume that full-rank LWE with parameters $(m, n, q, U_s, \chi_e)$ exist, which means that $n \leq m$. To differentiate full-rank LWE with possibly-not-full-rank LWE, we denote full-rank LWE with $\operatorname{LWE}^\ast$ and possibly-not-full-rank LWE.

@@

Since the secret is sampled from a uniformly random distribution, $\mathbf{s}_1$ and $\mathbf{s}_2$ are both uniformly random. Therefore, $\mathbf{s}_1 + B\mathbf{s}_2$ is also uniformly random. This means that $(A_1, \mathbf{b})$ is a valid sample from $\operatorname{LWE}^\ast(m, n_1, q, U_s, \chi_e)$ where $n_1 < n$. Assuming that there is a unique solution to $(A_1, \mathbf{b})$, if we can solve this instance of full-rank LWE, then the solution must be $\mathbf{s}^\prime$ such that $\mathbf{b} - A_1\mathbf{s}^\prime = \mathbf{e}$ is exactly the error term of the original $(A, \mathbf{b})$. With the error term recovered, the LWE problem $(A, \mathbf{b})$ becomes solving linear equations, which can be efficiently computed using Gaussian eliminations.

It remains to show that given an oracle that solves full-rank LWE of a higher dimension $n$, we can solve full-rank LWE of a lower dimension $n_1 < n$.

Given $(A_1, \mathbf{b})$ where $A \in \mathbb{Z}_q^{m \times n_1}$ is full-rank, I claim without proof that there exists an efficient algorithm that can output a vector $\mathbf{a}^\ast \in \mathbb{Z}_q^m$ such that $[A \;\vert\; \mathbf{a}]$ is still full-rank (see problem 1). Using this algorithm, we can augment $A_1$ into a full-rank matrix with $n$ columns $A^\prime = [ A_1 \;\vert\; A_2^\prime ]$. We can sample $\mathbf{s}_2^\prime$ from $U_s$ and compute $\mathbf{b}^\prime = \mathbf{b} + A_2^\prime\mathbf{s}_2^\prime$, then $(A^\prime, \mathbf{b}^\prime)$ is a valid instance of $\operatorname{LWE}^\ast(m, n, q, U_s, \chi_e)$. Assuming that $(A^\prime, \mathbf{b}^\prime)$ has unique solution, then the solution must be $\mathbf{s}_1 \;\vert\; \mathbf{s}_2^\prime$, and we can obtain $\mathbf{s}_1$, which is the solution to $(A_1, \mathbf{b})$. Thus we have solved full-rank LWE of a lower dimension.

P.S. the second reduction is very similar to the reduction in which an oracle for higher dimension LWE (not necessarily full-rank) can be used to solve lower dimension LWE by augmenting the lower dimension LWE instance. The only difference is that augmenting the lower dimension instance requires sampling linearly independent columns instead of sampling random columns.
